---
title: "Statistical Inference - Course Project 2"
date: "Tuesday, November 18 2014"
output: pdf_document
---
## Summary
In this project, we will analyze the ToothGrowth data in the R datasets package. The following are the objectives of this project: 

1. Load the ToothGrowth data and perform some basic exploratory data analyses 
2. Provide a basic summary of the data.
3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. 

## 1. Exploratory Data Analysis
We first load the datasets package and examine the structure of the ToothGrowth dataset. 

```{r}
library (datasets)
data(ToothGrowth)
str (ToothGrowth)
```

Now we look at some basic statistics of this dataset. 
```{r}
summary (ToothGrowth)
```

## 2. Summary and Graphical Analysis
Now we examine the data graphically. The plot below shows the `len` plotted against `dose` color conditioned on `supp`. We see that as the `dose` increases, the `len` increases as well. However, it is not clear whether the `supp` has any effect. 

```{r}
library (ggplot2)
par (mfrow = c(1,1))
ggplot (aes(x = dose, y = len), data = ToothGrowth) + 
    geom_point(aes (color = supp)) + 
    labs (list(title = "ToothGrowth by Dose and Supp", x = "Dose", y = "Supp"))
```

Now, we look at a boxplot of `supp` against `len` to check whether we can see any discernable difference or pattern. However, the plot shows that there is a lot of overlap between the two delivery methods i.e. `supp`. So we still do not know whether `supp` has a significant effect on `len` or not. 
```{r}
par (mfrow = c(1,1))
ggplot(aes(x=supp, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=supp)) + 
    labs (list(title = "ToothGrowth by Supp", x = "Supp", y = "Length"))
```

Now, we look at a boxplot of `dose` against `len` to check whether we can see any pattern. The plot shows that as `dose` increases, the `len` increases as well. It is possible that `dose` has a significant effect on `len`. 
```{r}
par (mfrow = c(1,1))
ggplot(aes(x=as.factor(dose), y=len), data=ToothGrowth) + geom_boxplot(aes(fill=dose)) + 
    labs (list(title = "ToothGrowth by Dose", x = "Dose", y = "Length"))
```

## 3. Confidence Intervals & Hypothesis Tests
First, we use the t-test to check whether the `supp` is statistically significant. 

```{r}
tTestSupp <- t.test(len ~ supp, data = ToothGrowth)
``` 
The p-value is `r tTestSupp$p.value`, which is relatively high. The 95% confidence interval contains zero. This indicates that we can not reject the null hypothesis that the two supplement types have no effect on tooth length `len`. 

We will now test whether the dose has a significant effect on the tooth length. We know that the `dose` variable has three distinct values. We cannot use a t-test to test for the significance of all three values. So we have to divide them into separate sub-datasets. Then we use t-tests to check whether the `dose` has significant effect on the tooth length `len`. 

```{r}
# Sub-groups based on dose pairs. 
dose051 <- subset (ToothGrowth, dose %in% c(0.5, 1.0)) 
dose052 <- subset (ToothGrowth, dose %in% c(0.5, 2.0)) 
dose12 <- subset (ToothGrowth, dose %in% c(1.0, 2.0)) 

# Check for differences in length due to change in dose levels: (0.5, 1.0)
t.test(len ~ dose, data = dose051)

# Check for differences in length due to change in dose levels: (0.5, 2.0)
t.test(len ~ dose, data = dose052)

# Check for differences in length due to change in dose levels: (1.0, 2.0)
t.test(len ~ dose, data = dose12)
```

In all three dose t-tests, the p-value is less than the fischer p-value of 0.05. Also, none of the 95% confidence interval contains zero. We see that the mean tooth length `len` increases with doseage `dose`. So we should reject the null hypothesis concluding that an increase in the dose level leads to an increase in tooth length.

## 4. Conslusions
1. Inreasing the doseage `dose` increases tooth growth `len`. 
2. The type of supplement used `supp`, whether its orange juice or vitamin C, has no discernable effect on tooth growth.

#### Assumptions: 

1. Variances between the different populations of guinea pigs are not the same. 
2. The populations are independent. 
3. Experiment was conducted by randomizing the sample guinea pig selection. 
